---
title: "eds213_data_cleaning_assign"
author: "Amanda Overbye"
format: html
editor: visual
---

# EDS-213 Data Cleaning Assignment

### Load libraries

```{r}
#| message: false  

# Libraries
library(tidyverse)
```

### Load data

```{r}
# file names

datadir_raw <- "data/raw/"

datadir_processed <- file.path("data", "processed/")

species_file <- "species_presence.csv"

snowsurvey_file <- "snow_cover.csv"

raw_snow_survey_file <- "ASDN_Snow_survey.csv"
```

```{r}
# Load data
snowsurvey_csv <- read_csv(file.path(datadir_processed, snowsurvey_file))
```

```{r}
# View dataset
glimpse(snowsurvey_csv)
```

# Clean Water_cover

```{r}
# Look for any suspicious values
snowsurvey_csv %>% 
  count(Water_cover)
```

```{r}
# Look for non-numeric values
snowsurvey_csv %>% 
  count(Water_cover) %>%
  filter(is.na(as.numeric(Water_cover)))
```

### Standardizing NAs

```{r}
# Turn n/a into NA
snowsurvey_fixed <- snowsurvey_csv %>% 
  mutate(Water_cover = ifelse(Water_cover == "n/a", NA, Water_cover))
```

```{r}
# Turn 'unk' (unknown) into NA
snowsurvey_fixed <- snowsurvey_fixed %>% 
  mutate(Water_cover = ifelse(Water_cover == "unk", NA, Water_cover))
```

```{r}
# Check results
snowsurvey_fixed %>% 
  count(Water_cover) %>%
  filter(is.na(as.numeric(Water_cover)))
```

### Turning Water_cover into a numeric column

I am deciding to turn Water_cover into a numeric column because it represents percentages out of 100.

```{r}
snowsurvey_fixed <- snowsurvey_fixed %>% 
  mutate(Water_cover = as.numeric(Water_cover))
```

### Looking for inappropriate numbers

Here I am looking for numbers that do not make sense in a column of percentages

```{r}
# Looking for values over 100%
snowsurvey_fixed %>% 
  filter(Water_cover > 100) 
```

```{r}
# Looking for values under 0
snowsurvey_fixed %>% 
  filter(Water_cover < 0) 
```

### Saving the csv

```{r}
# Saving changes as csv
write_csv(snowsurvey_fixed, file.path(datadir_processed, "water_cover_fixed.csv"))
```

# Clean Land_cover

```{r}
# Look for any suspicious values
snowsurvey_fixed %>% 
  count(Land_cover)
```


```{r}
# Looking for non-standardized NAs
snowsurvey_fixed %>% 
  count(Land_cover) %>%
  filter(is.na(as.numeric(Land_cover)))
```
### Standardizing NAs

```{r}
# Turn n/a into NA
snowsurvey_fixed <- snowsurvey_fixed %>% 
  mutate(Land_cover = ifelse(Land_cover == "n/a", NA, Land_cover))
```

```{r}
# Turn 'unk' (unknown) into NA
snowsurvey_fixed <- snowsurvey_fixed %>% 
  mutate(Land_cover = ifelse(Land_cover == "unk", NA, Land_cover))
```

```{r}
# Checking work
snowsurvey_fixed %>% 
  count(Land_cover) %>%
  filter(is.na(as.numeric(Land_cover)))
```
### Turning Land_cover into a numeric column

```{r}
snowsurvey_fixed <- snowsurvey_fixed %>% 
  mutate(Land_cover = as.numeric(Land_cover))
```

### Looking for outlier numbers

```{r}
# Looking for values over 100%
snowsurvey_fixed %>% 
  filter(Land_cover > 100) 
```

```{r}
# Looking for values under 0
snowsurvey_fixed %>% 
  filter(Land_cover < 0) 
```
```{r}
# Removing negative Land_cover values
snowsurvey_fixed <- snowsurvey_fixed %>% 
  mutate(Snow_cover = ifelse(Land_cover > 100, NA, Land_cover))
```

### Adding up the percentage points

```{r}
snowsurvey_clean <- snowsurvey_fixed %>%
  mutate(
    Land_cover = ifelse(Land_cover + Water_cover + Snow_cover != 100, NA, Land_cover),
    Water_cover = ifelse(Land_cover + Water_cover + Snow_cover != 100, NA, Water_cover),
    Snow_cover = ifelse(Land_cover + Water_cover + Snow_cover != 100, NA, Snow_cover)
  )
```

```{r}
# Look for any suspicious values
snowsurvey_fixed %>% 
  count(Snow_cover)
```

```{r}
# Look for any suspicious values
snowsurvey_clean %>% 
  count(Snow_cover)
```

